---
title: "MBXUCL package"
author: "Manon Martin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MBXUCL package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
require(knitr)
require(pander)
require(MBXUCL)


```

# Overview
This document provides a brief summary on how to use the MBXUCL library in R.

# Package installation

The library can be installed from a stand alone version contained in a personnal folder:
```{r install1, tidy=TRUE,  eval=FALSE}

install.packages(file.path("path to MBXUCL-package", "MBXUCL-package/MBXUCL"), repos = NULL, type="source")
require(MBXUCL)

```

Notes:

* Change the `path to MBXUCL-package`directory in `file.path` as it is in your computer
* Check for latest version of R and download additionnal required packages

Or it can be downloaded directly from Github:
```{r install2, tidy=TRUE,  eval=FALSE}
require(devtools)
install_github("ManonMartin/MBXUCL")
require(MBXUCL)
require(plyr)
```




# Available tools

## Inertia measures

```{r inertia}
data("HumanSerum")
Inertia.res = MBXUCL::Inertia(x = HumanSerumSpectra, y = ClassHS, print = TRUE)

kable(Inertia.res[["Between_within"]])

kable(Inertia.res[["Per_group"]])
```


## PCA
```{r pca}
data("HumanSerum")
PCA.res = MBXUCL::SVDforPCA(HumanSerumSpectra, ncomp=4)
```

### Eigenvalues

```{r eigPCA}
pander(PCA.res[["eigval"]][1:4])
```

### Scores

```{r Scores}
pander(PCA.res[["scores"]][1:10,])
```

Draw scores:

```{r scoreplot, out.width='70%', fig.width=10, fig.height=10}
DrawSL(PCA.res, type.obj = "PCA", drawNames=TRUE,
 createWindow=FALSE, main = "PCA score plot for HumanSerum dataset",
   class = as.character(ClassHS), axes =c(1,2), type.graph="scores")
```


### Loadings

```{r Loadings}
pander(PCA.res[["loadings"]][1:10,])
```

Draw loadings:
```{r, out.width='100%', fig.width=12, fig.height=8}
DrawSL(PCA.res, type.obj = "PCA", drawNames=TRUE,
 createWindow=FALSE, main = "PCA loadings plot for HumanSerum dataset",
    axes = c(1:2), type.graph ="loadings", loadingstype="l")
```





## Unsupervised clustering


### Unsupervised clustering on (GPL) intensities and associated MIC indexes


```{r ClustMIC, out.width='100%', fig.width=12, fig.height=12}
data("HumanSerum")
ClustMIC.res = MBXUCL::ClustMIC(Intensities = HumanSerumSpectra, nClust = 4, Trcl = ClassHS, Dendr = TRUE)

pander(ClustMIC.res)
```


### Unsupervised clustering on (GPL) binary presence/absence of peaks and associated MIC indexes


```{r binClustMIC, out.width='100%', fig.width=12, fig.height=12}
Pos = t(GPL[, seq(4,dim(GPL)[2], 2)])
Trcl = c(rep(1,8), rep(2,8), rep(3,8))
binClustMIC.res = MBXUCL::binClustMIC(Positions = Pos, Distance = "Jaccard", nClust = 3, Trcl, Dendr = TRUE)

pander(binClustMIC.res)
```

## PLS-DA 

```{r PLSDA, out.width='100%', fig.width=10, fig.height=10}
data("HumanSerum")
PLSDA.res = PLSDA(x = HumanSerumSpectra, y = ClassHS, nLV = NULL, drawRMSEP = TRUE)

perf.plsda = PLSDA.res[4:6]
pander(perf.plsda)
```

## OPLS-DA


```{r OPLSDA}

data("DataSimul")
x = DataSimul[["x"]]
y = DataSimul[["y"]]
m = dim(x)[2]
no=3
nb = 15
oplsda.res = OPLSDA(x=x, y=y, impT = FALSE,impG = FALSE, no=no, nb = nb)


```




Few graphical results :
  
  
  
* Choice of the number of orthogonal components

Eigenvalue criterion in Trygg (2001):
```{r , out.width='70%',fig.width=8, fig.height=8}
COL=rep("gray93",no)
mp=barplot(t(oplsda.res$CV),
           axes=F, axisnames=F, border=1,col=COL)
axis(1,at=mp, labels=c(1:no))
axis(2)
title(main="OPLS: Choice of the n[orthog. Components]", xlab="Orthogonal OPLS-DA Components",   ylab="||Wortho|| / ||p||", cex.main = 0.8)

```


Or by cross-validation:
  
  
```{r, out.width='70%', fig.width=10, fig.height=10, eval=TRUE}
cvOPLSDA.res = cvOPLSDA(x = x, y = y, k_fold = 10, NumOrtho = 10)
plot(1:length(cvOPLSDA.res$RMSECV), cvOPLSDA.res$RMSECV, main = "RMSECV", lty = 19)

```


* Scores and loadings plots:
  
```{r, out.width='100%', fig.width=12, fig.height=12}
Class = y
DrawSL(obj = oplsda.res, type.obj = "OPLSDA", drawNames = TRUE,
       createWindow = FALSE, main = NULL, class = Class, axes = c(1, 2),
       type.graph = "scores", loadingstype = "l",
       num.stacked = 4, xlab = NULL)


DrawSL(obj = oplsda.res, type.obj = "OPLSDA", drawNames = TRUE,
       createWindow = FALSE, main = NULL, axes = c(1, 2,3),
       type.graph = "loadings", loadingstype = "l",
       num.stacked = 4, xlab = NULL)

```


* S-plot


```{r, out.width='80%', fig.width=12, fig.height=12}

# S-plot
###########################
## correlation and covariance matrices
# covariance
s = as.matrix(x, ncol = ncol(x))
p1 = c()
for (i in 1:ncol(s)) {
  scov = cov(s[, i], oplsda.res$Tp)
  p1 = matrix(c(p1, scov), ncol = 1) # covariance x-T
}

# correlation
pcorr1 = c()
Tno=as.matrix(oplsda.res$Tp, ncol=1)
for (i in 1:nrow(p1)) {
  den = apply(Tno, 2, sd, na.rm = TRUE) * sd(s[, i])
  corr1 = p1[i, ]/den
  pcorr1 = matrix(c(pcorr1, corr1), ncol = 1) # correlation
}

# plot

plot(p1, pcorr1, xlab = "p(cov)[1]", ylab = "p(corr)[1]",
     main = "S-plot (OPLS-DA)", ylim=c(min(pcorr1, na.rm = T)*1.1,
                                       max(pcorr1, na.rm = T)*1.1), xlim=c(min(p1, na.rm = T)*1.1, 
                                                                           max(p1, na.rm = T)*1.1))

sel = p1*pcorr1
sel = order(sel, decreasing = TRUE)[1:nb]
text(p1[sel], pcorr1[sel], labels = colnames(s)[sel], cex = 0.7, pos = 1)
abline(v=0,lty = 2)
abline(h=0,lty = 2)



```



* Model coefficients

```{r, out.width='100%', out.width='70%', fig.width=12, fig.height=5}

delta=mean(sort(abs(oplsda.res$b.coef))[m-nb+c(0,1)])
xax = as.numeric(names(oplsda.res$b.coef))

par(mar=c(4,2,2,1))
plot(oplsda.res$b.coef,type="l",xaxt = "n", yaxt = "n", main="OPLS: Vector of descriptors' rank",xlab="ppm" )
abline(h=0)
abline(h=delta*c(-1,1),lty=2)
axis(side=2,cex.axis=0.7)
axis(side=1,at=c(1,seq(50,m,50)),
     labels=xax[c(1,seq(50,m,50))],cex.axis=0.7)

```


### Prediction of new observations

```{r}
xtrain = x[-c(1:5),]
ytrain = y[-c(1:5)]
xnew = x[c(1:5),]
oplsda.res = OPLSDA(x=xtrain, y=ytrain,
                    impT = FALSE,impG = FALSE, no=2, nb = 15, out.path = ".")
OPLSDA_pred(ropls = oplsda.res, x.new = xnew)


```
