---
title: "MBXUCL package"
author: "Manon Martin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MBXUCL package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
require(knitr)
require(pander)
require(MBXUCL)

```

# Overview
This document provides a brief summary on how to use the MBXUCL library in R.

# Package installation

The library can be installed from a stand alone version contained in a personnal folder:
```{r install1, tidy=TRUE,  eval=FALSE}

install.packages(file.path("path to MBXUCL-package", "MBXUCL-package/MBXUCL"), repos = NULL, type="source")
require(MBXUCL)

```

Notes:

* Change the `path to MBXUCL-package`directory in `file.path` as it is in your computer
* Check for latest version of R and download additionnal required packages

Or it can be downloaded directly from Github:
```{r install2, tidy=TRUE,  eval=FALSE}
require(devtools)
install_github("ManonMartin/MBXUCL")
require(MBXUCL)

```




# Available tools

## Inertia measures

```{r inertia}
data("HumanSerum")
Inertia.res = MBXUCL::Inertia(x = HumanSerumSpectra, y = ClassHS, print = TRUE)

kable(Inertia.res[["Between_within"]])

kable(Inertia.res[["Per_group"]])
```


## PCA
```{r pca}
data("HumanSerum")
PCA.res = MBXUCL::SVDforPCA(HumanSerumSpectra, ncomp=4)
```

### Eigenvalues

```{r eigPCA}
pander(PCA.res[["eigval"]][1:4])
```

### Scores

```{r Scores}
pander(PCA.res[["pcs"]][1:10,])
```

Draw scores:

```{r scoreplot, out.width='70%', fig.width=10, fig.height=10}
DrawPCA(PCA.res, drawNames=TRUE,
 createWindow=FALSE, main = "PCA score plot for HumanSerum dataset",
   class = as.character(ClassHS), axes =c(1,2), type ="scores")
```


### Loadings

```{r Loadings}
pander(PCA.res[["pcv"]][1:10,])
```

Draw loadings:
```{r, out.width='100%', fig.width=12, fig.height=8}
DrawPCA(PCA.res, drawNames=TRUE,
 createWindow=FALSE, main = "PCA loadings plot for HumanSerum dataset",
    axes = c(1:2), type ="loadings", loadingstype="l")
```





## Unsupervised clustering


### Unsupervised clustering on (GPL) intensities and associated MIC indexes


```{r ClustMIC, out.width='100%', fig.width=12, fig.height=12}
data("HumanSerum")
ClustMIC.res = MBXUCL::ClustMIC(Intensities = HumanSerumSpectra, nClust = 4, Trcl = ClassHS, Dendr = TRUE)

pander(ClustMIC.res)
```


### Unsupervised clustering on (GPL) binary presence/absence of peaks and associated MIC indexes


```{r binClustMIC, out.width='100%', fig.width=12, fig.height=12}
Pos = t(GPL[, seq(4,dim(GPL)[2], 2)])
Trcl = c(rep(1,8), rep(2,8), rep(3,8))
binClustMIC.res = MBXUCL::binClustMIC(Positions = Pos, Distance = "Jaccard", nClust = 3, Trcl, Dendr = TRUE)

pander(binClustMIC.res)
```

## PLS-DA 

```{r PLSDA, out.width='100%', fig.width=10, fig.height=10}
data("HumanSerum")
PLSDA.res = PLSDA(x = HumanSerumSpectra, y = ClassHS, nLV = NULL, drawRMSEP = TRUE)

perf.plsda = PLSDA.res[4:6]
pander(perf.plsda)
```


